\documentclass[11pt,letter, english,%twocolumn
]{article}
\pdfoutput=1

\usepackage[oldint]{custom_as}

\graphicspath{ {figures/} }

%%Drar in tabell och figurtexter
\usepackage[margin=10 pt]{caption}
%%För att lägga in 'att göra'-noteringar i texten
\usepackage{todonotes} %\todo{...}

%%För att själv bestämma marginalerna. 
\usepackage[
            top    = 2cm,
            bottom = 2cm,
            left   = 2cm, right  = 2cm
]{geometry}

\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\oh}{\ensuremath\mathpzc{o}}




\newcommand{\T}{{\!\mathsfrm{T}}}
\newcommand{\I}{{\mathbfrm{I}}}
\newcommand{\PC}{P_{\text{C}}}
\newcommand{\PN}{P_{\text{N}}}



\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{AMATH\,732}
\chead{Swings and Biorythms}
\rhead{Avery, Sundstr{\"o}m}

\usepackage{indentfirst}

\begin{document}

%%%%%%%%%%%%%%%%% vvv Inbyggd titelsida vvv %%%%%%%%%%%%%%%%%

\title{\vspace{-1cm}
Swings and Biorythms --- Parametrically forced oscillators}
\author{Leon Avery \and Andréas Sundstr{\"o}m}
\date{November 21, 2016}

\maketitle

\begin{abstract}\noindent
This report presents two different scenarios, which both can be
modeled as parametrically forced oscillators: pupming a playgroud
swing, and keeping the circadian rythm in sync with the
environment. 
Children's swings and biological clocks are very different, but they
present a common problem. Each is a natural oscillator with an
intrinsic free-running rhythm. And each is driven by a forcing
function that works, not by directly pushing the oscillator itself,
but by periodically modifying the parameters of the oscillator.
The swing is modeled as a point-mass pendulum with a time
dependent length. We investigate conditions needed to pump the
swing. Under those conditions we show that the amplitude, to a first
approximation, will increas exponentially.
In the case of the biological clock, the forcing is external light,
which modifies the rate of biochemical reactions such as the synthesis
of messenger RNA. \todo{Maybe some results on biorythms}
Mathematically, both of these problems are
united by Floquet Theory: a general analysis of linear differential
equations with periodic parametric forcing. Although Floquet Theory
does not directly give solutions to the differential equations, it
sharply constrains the form these solutions can take. In particular,
it helps to place bounds on the frequencies that will effectively pump
a swing, and it predicts the existence of periodic solutions to the
phase gradient differential equation for the circadian clock.  

% In the case of the circadian rythms, we study a model that allows for
% light to affect the biological clock so that it will stay in phase
% with the day/night cycle.
% \todo[inline]{Something on the results of the biorythms.}

% Children's swings and biological clocks are very different, but they
% present a common problem. Each is a natural oscillator with an
% intrinsic free-running rhythm. And each is driven by a forcing
% function that works, not by directly pushing the oscillator itself,
% but by periodically modifying the parameters of the oscillator. In the
% case of the swing, the forcing is the up-and-down motion of the child,
% which modifies the effective length of the swing. In the case of the
% biological clock, the forcing is external light, which modifies the
% rate of biochemical reactions such as the synthesis of messenger
% RNA. Both oscillators have in part a common goal: to match the
% frequency of the oscillation to that frequency of the forcing. For the
% swing, this matching means that the child's motions occur at a
% consistent point in the cycle --- for instance, the child consistently
% moves upward when the swing is at its lowest point. This
% synchronization of forcing and the swing cycle allows the child to
% most effectively pump energy into the swing's motion. For the clock,
% this matching (called entrainment) ensures a match between the state
% of biochemical quantities (e.g. high PERIOD protein levels) and
% specific points in the day/night cycle, (e.g. dawn), allowing the
% organism to anticipate events. Mathematically, these problems are
% united by Floquet Theory: a general analysis of linear differential
% equations with periodic parametric forcing. Although Floquet Theory
% does not directly give solutions to the differential equations, it
% sharply constrains the form these solutions can take. In particular,
% it helps to place bounds on the frequencies that will effectively pump
% a swing, and it predicts the existence of periodic solutions to the
% phase gradient differential equation for the circadian clock.  
\end{abstract}

%%%%%%%%%%%%%%%%% ^^^ Inbyggd titelsida ^^^ %%%%%%%%%%%%%%%%%

\input{Swing.input.tex}

\input{Floquet_theory.input.tex}

\input{Circadian_rythms.input.tex}


\section*{Bibliography}
\noindent
Burns, J.A. (1970). \textit{More on Pumping a Swing}. Am. J. Phys. \textbf{38}, 920-2

\bigskip\noindent
Ko, C.H., and Takahashi, J.S. (2006). \textit{Molecular components of the
mammalian circadian clock}. Hum. Mol. Genet. R271-7.

\bigskip\noindent
Konopka, R.J., and Benzer, S. (1971). \textit{Clock Mutants of Drosophila
melanogaster}. \textbf{68}, 2112–2116.

\bigskip\noindent
Partch, C.L., Green, C.B., Takahashi (2014).
\textit{Molecular architecture of the mammalian circadian clock}. Trends Cell
Biol. \textbf{24}, 90–99.

\bigskip\noindent
Pfeuty, B., Thommen, Q., and Lefranc, M. (2011). \textit{Robust Entrainment of
Circadian Oscillators Requires Specific Phase Response Curves}.

\bigskip\noindent
Rosato, E., Tauber, E., and Kyriacou, C.P. (2006). \textit{Molecular genetics
of the fruit-fly circadian clock}. Eur. J. Hum. Genet. \textbf{14},
729–738.

\bigskip\noindent
Strub, D.C. (2009). \textit{How do children swing?}  M.Eng. thesis, University
of Bristol, Dept. of Eng. Math.%ematics

\bigskip\noindent
Tea, P.L. Jr. and Falk, H. (1968). \textit{Pumping on a Swing}. Am. J. Phys. 
\textbf{36}, 1165-6

\bigskip\noindent
Ward, M.J. \textit{Basic Floquet Theory}
(http://www.math.ubc.ca/\~{}ward/teaching/m605/every2\_floquet1.pdf).


\end{document}





%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "Report_swing_circadian"
%%% End: 
