\documentclass[11pt,a4paper, 
swedish, english %% Make sure to put the main language last!
]{article}
\pdfoutput=1

%% Andréas's custom package 
%% (Will work for most purposes, but is mainly focused on physics.)
\usepackage{../custom_as}

%% Figures can now be put in a folder: 
\graphicspath{ {figures/} %{some_folder_name/}
}

%% If you want to change the margins for just the captions
\usepackage[margin=10 pt]{caption}

%% To add todo-notes in the pdf
\usepackage[%disable  %%this will hide all notes
]{todonotes} 

%% Change the margin in the documents
\usepackage[
%            top    = 3cm,              %% top margin
%            bottom = 3cm,              %% bottom margin
%            left   = 2.5cm, right  = 2.5cm %% left and right margins
]{geometry}


%% If you want to change the formatting of the section headers
%\renewcommand{\thesection}{...}

\newcommand{\EM}{\text{EM}}
\newcommand{\E}{\text{E}}
%\newcommand{\GR}{\text{GR}}
%\newcommand{\SR}{\text{SR}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}%% v v v v v v v v v v v v v v v v v v v v v v v v v v
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%% If you want to use an external file for the title page
%\input{titlepages.tex}


%%%%%%%%%%%%%%%%%%%% vvv Internal title page vvv %%%%%%%%%%%%%%%%%%%%%
\title{Gravitation and Cosmology -- FFM 071
\\ {\Large Hand-in set 2} }
\author{Andréas Sundström}
\date\today%{2018-02-13}

\maketitle

%%%%%%%%%%%%%%%%%%%% ^^^ Internal title page ^^^ %%%%%%%%%%%%%%%%%%%%%
%% If you want a list of all todos
%\todolist

\renewcommand{\thesubsection}{\arabic{section} (\roman{subsection})}
\setcounter{section}{8}

\section{3D anti-de Sitter}

The $\text{AdS}_3$ metric
\begin{equation}\label{eq9:metric1}
\rd{s}^2=-\rd{u}^2-\rd{v}^2+\rd{x}^2+\rd{y}^2
\end{equation}
under the constraint
\begin{equation}\label{eq9:constraint}
-u^2-v^2+x^2+y^2=-R^2,
\end{equation}
can be rewritten using the change of variables
\begin{equation}\label{eq9:var-change}
\begin{aligned}
&u=\sqrt{R^2+r^2}\cos(t/R)
\qc\qquad &x=r\cos\phi, \\
&v=\sqrt{R^2+r^2}\sin(t/R)
\qc\qquad &y=r\sin\phi.
\end{aligned}
\end{equation}
Firstly it is easy to confirm that \eqref{eq9:var-change} satisfy
\eqref{eq9:constraint} by noting that
\begin{equation}
u^2+v^2=R^2+r^2\qand
x^2+y^2=r^2.
\end{equation}
Then to calculate the metric, we need
\begin{equation}
\rd{u}=\pdv{u}{r}\rd{r}+\pdv{u}{t}\rd{t}+\pdv{u}{\phi}\rd\phi
=\frac{r\cos(t/R)\rd{r}}{\sqrt{R^2+r^2}}
-\frac{\sqrt{R^2+r^2}}{R}\sin(t/R)\rd{t}
\end{equation}
and similarly for
\begin{equation}
\rd{v}=\frac{r\sin(t/R)\rd{r}}{\sqrt{R^2+r^2}}
+\frac{\sqrt{R^2+r^2}}{R}\cos(t/R)\rd{t},
\end{equation}
which leads to
\begin{equation}
\rd{u}^2+\rd{v}^2=
\frac{r^2\rd{r}^2}{R^2+r^2}+\frac{R^2+r^2}{R^2}\rd{t}^2
\end{equation}
Then since we already recognize the polar coordinates for $x$ and $y$,
we can write down
\begin{equation}
\rd{x}^2+\rd{y}^2=\rd{r}^2+r^2\rd\phi^2.
\end{equation}
We can now write down the metric \eqref{eq9:metric1} in terms of the
new coordinates
\begin{equation}
\begin{aligned}
\rd{s}^2=&-\frac{R^2+r^2}{R^2}\rd{t}^2
+\qty[1-\frac{r^2}{R^2+r^2}]\rd{r}^2+r^2\rd\phi^2\\
=&-\frac{R^2+r^2}{R^2}\rd{t}^2
+\frac{R^2}{R^2+r^2}\rd{r}^2+r^2\rd\phi^2
\end{aligned}
\end{equation}

The proper distance from $r=0$ to $r=\infty$ can easily be calculated
as
\begin{equation}
S=\int_{r=0}^{r=\infty}\rd{s}=
\begin{Bmatrix}
\rd{t}=0\\
\rd{\phi}=0
\end{Bmatrix}
=\int_0^\infty\rd{r}
\overbrace{\frac{R}{\sqrt{R^2+r^2}}}^{\sim R/r\text{ for }r\gg R}
 =\infty.
\end{equation}
Meanwhile the time it takes for a light signal to traverses the same
path is
\begin{equation}
T=\int_{r=0}^{r=\infty}\rd{t}
=\int_0^\infty\rd{r}\frac{R^2}{R^2+r}
=\frac{\pi R}{2},
\end{equation}
where $\rd{t}$ have been calculated using
\begin{equation}
0=\rd{\tau^2}=\frac{R^2+r^2}{R^2}\rd{t}^2
-\frac{R^2}{R^2+r^2}\rd{r}^2
\end{equation}
for light.



\section{Maximally symmetric metric?}
We have a metric given by
\begin{equation}\label{eq10:ds}
\rd{s}^2=-\rd{t}^2+\rd{x}^2+\rd{y}^2+\rd{z}^2
-4\cosh(\frac{x}{2})
\qty[\cosh(\frac{x}{2})(\rd{t}+\rd{x})-\sinh(\frac{x}{2})\rd{y}]\rd{x},
\end{equation}
which in matrix form gives
\begin{equation}
g_{\mu\nu}=
\begin{bmatrix}
-1 & -(1+\cosh x) &0 &0\\
-(1+\cosh x) &-(1+2\cosh x) &+\sinh x&0\\
0&+\sinh x &1&0\\
0&0&0&1
\end{bmatrix}.
\end{equation}
Here we have used the hyperbolic identities
\begin{equation}
2\cosh^2\qty(\frac{x}{2})\equiv 1+\cosh x
\qand
2\cosh(\frac{x}{2})\sinh(\frac{x}{2})
\equiv \sinh x.
\end{equation}

To see if this is maximally symmetric, we calculate the Riemann
tensor. To do that we need the affine connection
\begin{equation}
\Gamma^\rho_{\mu\nu}=\frac{1}{2}g^{\rho\sigma}
\qty(\pd_{\mu}g_{\nu\sigma}+\pd_{\nu}g_{\mu\sigma}-\pd_{\sigma}g_{\mu\nu}).
\end{equation}
However we inside the parenthesis only the elements with $\mu\nu=xx$
survive. To prove this we take $\mu\nu=x\alpha$, where $\alpha\neq x$,
this gives
\begin{equation}
\pd_{x}g_{\alpha\sigma}+\cancelto{0}{\pd_{\alpha}}g_{x\sigma}-\pd_{\sigma}g_{x\alpha}
=\pd_{x}g_{\alpha\sigma}-\pd_{\sigma}g_{x\alpha}.
\end{equation}
The only way for either of the two remaining terms to survive is if
$\sigma=x$, but in that case the two terms are identical and cancel
each other. Then for $\mu\nu=\alpha\beta$, where $\alpha,\beta\neq x$,
we get
\vspace{-1em}
\begin{equation}
\cancelto{0}{\pd_{\alpha}}g_{\beta\sigma}
+\cancelto{0}{\pd_{\beta}}g_{\alpha\sigma}
-\cancel{\pd_{\sigma}g_{\alpha\beta}}=0.
\end{equation}
The last term cancel because the only non-constant elements in
$g_{\mu\nu}$ are on either $\mu=x$ or $\nu=x$, which is not the case
for $\mu\nu=\alpha\beta$. Hence the only non-vanishing elements of the
affine connection are $\Gamma^\rho_{xx}$. This then means that the
Riemann tensor,
\begin{equation}
{R^\rho}_{\sigma\mu\nu}=
\pd_{\mu}\Gamma^\rho_{\sigma\nu}-\pd_{\nu}\Gamma^\rho_{\sigma\mu}
+\Gamma^\rho_{\mu\tau}\Gamma^\tau_{\sigma\nu}
-\Gamma^\rho_{\nu\tau}\Gamma^\tau_{\sigma\mu},
\end{equation}
actually vanishes. To see this more clearly, we note that the
anti-symmetry in $\mu\nu$ means that $\mu\neq\nu$, but due to the
restriction to $\Gamma^\rho_{xx}$ we must have one of $\mu$ or $\nu$
to be $x$. Without loss of generality, we choose $\mu=x$ and
$\nu=\alpha\neq x$, and we get
\begin{equation}
{R^\rho}_{\sigma x\alpha}=
\pd_{x}\cancelto{0}{\Gamma^\rho_{\sigma\alpha}}
-\cancelto{0}{\pd_{\alpha}}\Gamma^\rho_{\sigma x}
+\Gamma^\rho_{x\tau}\cancelto{0}{\Gamma^\tau_{\sigma\alpha}}
-\cancelto{0}{\Gamma^\rho_{\alpha\tau}}\Gamma^\tau_{\sigma x}=0.
\end{equation}
This is in other words a flat space-time, and is therefore maximally
symmetric.
% Note to self: See Weinberg p. 380:
% "Thus any flat metric admits $N(N+1)/2$ independent Killing vectors,
% and is therefore maximally symmetric."


We will therefore now try to get \eqref{eq10:ds} into the familiar
form of the Minkowski metric,
\begin{equation}\label{eq10:Mink1}
\rd{s'}^2=-\rd{t'}^2+\rd{x'}^2+\rd{y'}^2+\rd{z'}^2.
\end{equation}
To do this we introduce the new variables $t'=t'(t,x)$, $x'=x'(x)$,
$y'=y'(x,y)$, and $z'=z'(z)=z$. For $z'$ the choice is obvious, and
for the other three variables it is relatively easy to convince
yourself that that they should only depend on $x$ and their non-primed
counterpart. WE can now write out the metric \eqref{eq10:Mink1}
(ignoring $z'$ and $z$ since we already know that transformation) 
\begin{equation}%\label{eq10:Mink2}
\begin{aligned}
\rd{s'}^2=&
-\qty(\pdv{t'}{t})^2\rd{t}^2-2\pdv{t'}{t}\pdv{t'}{x}\id{t}\id{x}
-\qty(\pdv{t'}{x})^2\rd{x}^2\\
&+\qty(\pdv{x'}{x})^2\rd{x}^2\\
&+\qty(\pdv{y'}{x})^2\rd{x}^2+2\pdv{y'}{x}\pdv{y'}{y}\id{x}\id{y}
+\qty(\pdv{y'}{y})^2\rd{y}^2.
\end{aligned}
\end{equation}
For this to equal the original metric \eqref{eq10:ds}, we must have
\begin{equation}\label{eq10:cond}
\begin{cases}
\qty(\pdv{t'}{t})^2=1\\
\qty(\pdv{y'}{y})^2=1\\
\qty(\pdv{x'}{x})^2-\qty(\pdv{t'}{x})^2+\qty(\pdv{y'}{x})^2
=-(1+2\cosh x)\\
\pdv{t'}{t}\pdv{t'}{x}=1+\cosh x\\
\pdv{y'}{x}\pdv{y'}{y}=\sinh x.
\end{cases}
\end{equation}
If we choose the positive solutions to the first two equations, we get
\begin{equation}
\begin{cases}
\qty(\pdv{t'}{t})=+1\\
\qty(\pdv{y'}{y})=+1\\
\end{cases}
\quad\Longrightarrow\quad
\begin{cases}
t'(t,x)=t+f(x)\\
y'(x,y)=y+g(x),
\end{cases}
\end{equation}
which means that the last two equations in \eqref{eq10:cond} become
\begin{equation}
\begin{cases}
\dv{f}{x}=1+\cosh x\\
\dv{g}{x}=\sinh x
\end{cases}
\quad\Longrightarrow\quad
\begin{cases}
t'(t,x)=t+x+\sinh x\\
y'(x,y)=y+\cosh x.
\end{cases}
\end{equation}
It is now straight forward to show that the middle equation in
\eqref{eq10:cond} becomes
\begin{equation}
\qty(\pdv{x'}{x})^2
\cancel{-(1+2\cosh x)}
\overbrace{-\cosh^2 x+\sinh^2x}^{=-1}
=\cancel{-(1+2\cosh x)},
\end{equation}
which means that we also here can choose $\pdv*{x'}{x}=+1$. Every
thing is self-consistent, meaning that the full transformation of
variables 
\begin{equation}
\begin{cases}
t'=t'(t,x)=t+x+\sinh x\\
x'=x'(x)=x\\
y'=y'(x,y)=y+\cosh x\\
z'=z'(z)=z
\end{cases}
\end{equation}
results in $\rd{s'}^2$ from \eqref{eq10:Mink1} and $\rd{s}^2$ from
\eqref{eq10:ds} being equal.



\section{The non-flat torus}
On the non-flat torus embedded in three dimensions 
\begin{equation}
\begin{cases}
x=r_2A(\phi_2)\cos\phi_1\\
y=r_2A(\phi_2)\sin\phi_1\\
z=r_2\sin\phi_2
\end{cases}
\qq{with}
A(\phi_2)=r_1+r_2\cos\phi_2,
\end{equation}
and the metric, given by 
\begin{equation}
\rd{s}^2=\rd{x}^2+\rd{y}^2+\rd{z}^2
=A^2(\phi_2)\rd\phi_1^2+{r_2}^2\rd\phi_2^2,
\end{equation}
is
\begin{equation}
g_{\mu\nu}=
\begin{bmatrix}
A^2(\phi_2)&0\\
0&{r_2}^2
\end{bmatrix}.
\end{equation}
From this it is a simple task to compute the Affine connection
\begin{equation}
\Gamma^\rho_{\mu\nu}=\frac{1}{2}g^{\rho\sigma}
\qty(\pd_{\mu}g_{\nu\sigma}+\pd_{\nu}g_{\mu\sigma}-\pd_{\sigma}g_{\mu\nu}),
\end{equation}
whose only non-zero elements are 
\begin{equation}\label{eq11:affine}
\begin{aligned}
\Gamma^1_{12}=\Gamma^1_{21}
=&\frac{1}{2}g^{11}\pd_2g_{11}
=-\frac{r_2\sin\phi_2}{A(\phi_2)}\\
\Gamma^2_{11}=&-\frac{1}{2}g^{22}\pd_2g_{11}
=\frac{A(\phi_2)\sin\phi_2}{r_2}.
\end{aligned}
\end{equation}
Here we have used $\pd_2A^2(\phi_2)=-2A(\phi_2){r_2}\sin\phi_2$. Then
the Riemann tensor,
\begin{equation}
{R^\rho}_{\sigma\mu\nu}=
\pd_{\mu}\Gamma^\rho_{\sigma\nu}-\pd_{\nu}\Gamma^\rho_{\sigma\mu}
+\Gamma^\rho_{\mu\tau}\Gamma^\tau_{\sigma\nu}
-\Gamma^\rho_{\nu\tau}\Gamma^\tau_{\sigma\mu},
\end{equation}
is also seen to only have these few non-zero elements
\begin{equation}
\begin{aligned}
{R^1}_{212}=-{R^1}_{221}=&-\pd\Gamma^1_{21}-\Gamma^1_{21}\Gamma^1_{21}
=\frac{r_2}{A(\phi_2)}\cos\phi_2,\\
{R^2}_{112}=-{R^1}_{121}=&-\pd\Gamma^2_{11}+\Gamma^2_{11}\Gamma^1_{21}
=-\frac{A(\phi_2)}{r_2}\cos\phi_2.
\end{aligned}
\end{equation}
This gives the Ricci tensor
\begin{equation}
R_{\mu\nu}={R^\rho}_{\mu\rho\nu}={R^1}_{\mu1\nu}+{R^2}_{\mu2\nu}
=\begin{bmatrix}
\frac{A(\phi_2)}{r_2}\cos\phi_2&0\\
0&\frac{r_2}{A(\phi_2)}\cos\phi_2
\end{bmatrix},
\end{equation}
and the Ricci scalar
\begin{equation}
R=g^{\mu\nu}R_{\mu\nu}=g^{11}R_{11}+g^{22}R_{22}
=\frac{2\cos\phi_2}{r_2A(\phi_2)}
=\frac{2\cos\phi_2}{r_2(r_1+r_2\cos\phi_2)}.
\end{equation}
We now also note that $R_{\mu\nu}=g_{\mu\nu}R/2$ as is the case for
maximally symmetric spaces.

\todo[inline]{Still need to calculate the Killing vectors. Can't seem
  to separate the varaibles properly.}



\section{Wormholes}
Here we have the metric
\begin{equation}\label{eq12:ds}
\rd{s}^2=-\rd{t}^2+\rd{r}^2
+\qty(r^2+a^2)\qty(\rd\theta^2+\sin^2\theta\id\phi^2),
\end{equation}
giving 
\begin{equation}
g_{\mu\nu}=\text{Diag}
\begin{bmatrix}
-1,& 1,& (r^2+a^2),& (r^2+a^2)\sin^2
\end{bmatrix}.
\end{equation}
We now want to find the geodesic equations
\begin{equation}
\dv[2]{x^\mu}{\tau}+\Gamma^\mu_{\rho\nu}\dv{x^\rho}{\tau}\dv{x^\nu}{\tau}=0. 
\end{equation}
To do that we need the affine connection which, since $g_{\mu\nu}$ is
diagonal and only depends on $r$ and $\theta$, is relatively straight
forward to compute. Its only non-zero elements are
\begin{equation}\label{eq12:Gamma}
\begin{aligned}
&\Gamma^r_{\theta\theta}=-r\qc
\Gamma^r_{\phi\phi}=-r\sin^2\theta\\
&\Gamma^{\theta}_{r\theta}=\Gamma^{\theta}_{\theta r}
=\frac{r}{r^2+a^2}\qc
\Gamma^{\theta}_{\phi\phi}=-\sin\theta\,\cos\theta\\
&\Gamma^{\phi}_{r\phi}=\Gamma^{\phi}_{\phi r}=\frac{r}{r^2+a^2}\qc
\Gamma^\phi_{\theta\phi}=\Gamma^\phi_{\phi\theta}=\cot\theta,
\end{aligned}
\end{equation}
which gives the geodesic equations
\begin{equation}
\begin{cases}
\dv[2]{t}{\tau}=0\\
\dv[2]{r}{\tau}-r\qty(\dv{\theta}{\tau})^2
-r\sin^2\theta\,\qty(\dv{\phi}{\tau})^2=0\\
\dv[2]{\theta}{\tau}+\frac{2r}{r^2+a^2}\dv{r}{\tau}\dv{\theta}{\tau}
-\sin\theta\cos\theta\qty(\dv{\phi}{\tau})^2=0\\
\dv[2]{\phi}{\tau}+\frac{2r}{r^2+a^2}\dv{r}{\tau}\dv{\phi}{\tau}
+\cot\theta\dv{\phi}{\tau}\dv{\theta}{\tau}=0.
\end{cases}
\end{equation}
The last two equations can be rewritten, by multiplying with
$(r^2+a^2)$ (and also with $\sin^2\theta$ for the last equation), to
\begin{equation}
\begin{cases}
\dv{\tau}\qty[\qty(r^2+a^2)\dv{\theta}{\tau}]
=\qty(r^2+a^2)\sin\theta\cos\theta\qty(\dv{\phi}{\tau})^2\\[0.5ex]
\dv{\tau}\qty[\qty(r^2+a^2)\sin^2\theta\dv{\phi}{\tau}]=0.
\end{cases}
\end{equation}

\todo[inline]{Find wormhole solution!}

Using \eqref{eq12:Gamma} we can easily calculate the Riemann tensor
whose non-zero elements are
\begin{equation}
\begin{aligned}
&{R^r}_{\theta r\theta}=-{R^r}_{\theta\theta r}
=-\frac{a^2}{r^2+a^2} \qc
{R^r}_{\phi r\phi}=-{R^r}_{\phi\phi r}
=-\frac{a^2\sin^2\theta}{R^2+a^2}\\
&{R^\theta}_{r\theta r}=-{R^\theta}_{rr\theta}
=-\frac{a^2}{(r^2+a^2)^2}\qc
{R^\theta}_{\phi\theta\phi}=-{R^\theta}_{\phi\phi\theta}
=\frac{a^2\sin^2\theta}{r^2+a^2},\\
&{R^\phi}_{r\phi r}=-{R^\phi}_{rr\phi}
=-\frac{a^2}{(r^2+a^2)^2}\qc
{R^\phi}_{\theta\phi\theta}=-{R^\phi}_{\theta\theta\phi}
=\frac{a^2}{r^2+a^2}.
\end{aligned}
\end{equation}
From this we can compute the Ricci tensor and scalar. It is clear from
the Riemann tensor above that $R_{\mu\nu}={R^\rho}_{\mu\rho\nu}$ must
be diagonal and is given by
\begin{equation}
R_{\mu\nu}=\text{Diag}
\begin{bmatrix}
0,& -\frac{2a^2}{(r^2+a^2)^2},& 0, &0
\end{bmatrix},
\end{equation}
and then the curvature tensor trivially becomes
\begin{equation}
R=g^{\mu\nu}R_{\mu\nu}=-\frac{2a^2}{(r^2+a^2)^2}.
\end{equation}
We also see that $a\to0$ recovers a flat space, as we would expect
from the metric.

We now can write down the Einstein tensor as
\begin{equation}
G_{\mu\nu}=R_{\mu\nu}-\frac{R}{2}g_{\mu\nu}
=\frac{a^2}{(r^2+a^2)^2}\,\text{Diag}\!
\begin{bmatrix}
-1,& -1,&
(r^2+a^2),&
(r^2+a^2)\sin^2\theta
\end{bmatrix}.
\end{equation}
There are however problems with this Einstein tensor. According to
Einstein's equations 
\begin{equation}
G_{\mu\nu}=8\pi G T_{\mu\nu},
\end{equation}
and therefore since $G_{00}<0$ also $T_{00}<0$, but $T_{00}$
represents the energy density which can not be negative. 

\todo[inline]{Find all isometries!}





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}%% ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^ ^
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  contravariant Stereographic covariant variational
%  LocalWords:  azimuthal affine parametrization Ricci Minkowski
